What's in a Name? Linear Temporal Logic Literally Represents Time Lines

### 2023-01-02 paper/main.tex
typo fix
getting rid of bang

\begin{example}
    In \cref{fig:ex15}, one can reason about one timeline: the atomic proposition $a$ is false for finitely many time steps as signified by two nodes with $!a$ separated by $\cdots$ (could also be zero time step); followed by one node with $a$ which substantiates the specification of ``$a$ eventually true''; once $a$ is true at some point, the later time steps can behave arbitrarily as signified by the infinite run of $1$ (true) in the grey box.
    \begin{figure}[!h]
        \centering
        \includegraphics[scale=0.4]{examples/ex15/ex15.png}
        \caption{Example timeline for specification ``$a$ is eventually true''}
        \label{fig:ex15}
    \end{figure}
\end{example}

@@

\begin{example}
    The formula $\Phi = \always (a \limplies \eventually (\neg a))$ generates un-simplified $\omega$-regular expression $((!a) | (aa^*(!a)))((!a) | (aa^*(!a)))^{\omega}$, and simplified version $((!a) | (aa^*(!a)))^{\omega}$, which corresponds to the two timelines in \cref{fig:unsimplified} and \cref{fig:simplified}.
    \begin{figure}[h!]
        \centering
                \resizebox{\textwidth}{!}{
        \begin{tikzpicture}
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
        \node (m) at ($(a1)!0.5!(a2)$) {};
        \node[shape=rectangle, fill=gray!30!white, style={minimum width=10.9cm, minimum height=3cm}] (g) [right= -1.1cm of m] {};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a3) [right= 0.75cm of a2] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a4) [right= 0.75cm of a3] {\!\!\!$\dots$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a5) [right= 0.75cm of a4] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a6) [right= 0.75cm of a5] {\!\!\!$\lnot a$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) [left= 12.5cm of m] {\!\!\!start\!\!\!};

@@

        \end{tikzpicture}}
%        \includegraphics[scale=0.3]{examples/ex9/ex9-unsimplified.png}
        \caption{Timeline for un-simplified $\omega$-regular expression $((!a) | (aa^*(!a)))((!a) | (aa^*(!a)))^{\omega}$}
        \label{fig:unsimplified}
    \end{figure}
    \begin{figure}[h!]
        \centering
        \resizebox{.8\textwidth}{!}{
        \begin{tikzpicture}
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
  %      \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p3) [left= 1cm of p2] {\!\!\!$\lnot p$\!\!\!};
        \node (m) at ($(a1)!0.5!(a2)$) {};
        \node[shape=rectangle, fill=gray!30!white, style={minimum width=10.9cm, minimum height=3cm}] (g) [right= -1.1cm of m] {};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a3) [right= 0.75cm of a2] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a4) [right= 0.75cm of a3] {\!\!\!$\dots$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a5) [right= 0.75cm of a4] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a6) [right= 0.75cm of a5] {\!\!\!$\lnot a$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) [left= 2cm of m] {\!\!\!start\!\!\!};
        \draw[thick, line width=1.25] (s) edge[-latex] (a1) (s) edge[-latex] (a2) (a2) edge[-latex] (a3) (a3) edge[-latex] (a4) (a4) edge[-latex] (a5) (a5) edge[-latex] (a6);
        \end{tikzpicture}}
  %      \includegraphics[scale=0.3]{examples/ex9/ex9.png}
        \caption{Timeline for simplified $\omega$-regular expression $((!a) | (aa^*(!a)))^{\omega}$}
        \label{fig:simplified}
    \end{figure}
\end{example}

@@

\begin{definition}[Timeline Length]
    We define the timeline length of a regular expression $\tllen(A)$ recursively as follows
    \begin{align*}
        \tllen(\emptyset)   &= -\infty\\
        \tllen(\epsilon)        &= 0\\
        \tllen(a)           &= 1\\
        \tllen(A_1 A_2)     &= \tllen(A_1) + \tllen(A_2)\\
        \tllen(A_1 + A_2)   &= \max(\tllen(A_1, A_2))\\
        \tllen(A^*)         &= 2\tllen(A) + 1
    \end{align*}
    We also extend this definition to $\omega$-regular expressions as follows
    \begin{align*}
        \tllen(A^\omega)    &= \tllen(A)\\
        \tllen(AB)          &= \tllen(A) + \tllen(B)\\
        \tllen(B_1 + B_2)   &= \max(\tllen(B_1), \tllen(B_2))
    \end{align*}
    This quantity $\tllen$ captures the length of the longest path in our timeline graph visualization.
\end{definition}

>>>>

\begin{example}
    In \cref{fig:ex15}, one can reason about one timeline: the atomic proposition $a$ is false for finitely many time steps as signified by two nodes with $\lnot a$ separated by $\cdots$ (could also be zero time step); followed by one node with $a$ which substantiates the specification of ``$a$ eventually true''; once $a$ is true at some point, the later time steps can behave arbitrarily as signified by the infinite run of $1$ (true) in the grey box.
    \begin{figure}[!h]
        \centering
        \includegraphics[scale=0.4]{examples/ex15/ex15.png}
        \caption{Example timeline for specification ``$a$ is eventually true''}
        \label{fig:ex15}
    \end{figure}
\end{example}

@@

\begin{example}
    The formula $\Phi = \always (a \limplies \eventually (\neg a))$ generates un-simplified $\omega$-regular expression $((\lnot a) | (aa^*(\lnot a)))((\lnot a) | (aa^*(\lnot a)))^{\omega}$, and simplified version $((\lnot a) | (aa^*(\lnot a)))^{\omega}$, which corresponds to the two timelines in \cref{fig:unsimplified} and \cref{fig:simplified}.
    \begin{figure}[h!]
        \centering
                \resizebox{\textwidth}{!}{
        \begin{tikzpicture}
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
        \node (m) at ($(a1)!0.5!(a2)$) {};
        \node[shape=rectangle, fill=gray!30!white, style={minimum width=10.9cm, minimum height=3cm}] (g) [right= -1.1cm of m] {};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a3) [right= 0.75cm of a2] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a4) [right= 0.75cm of a3] {\!\!\!$\dots$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a5) [right= 0.75cm of a4] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a6) [right= 0.75cm of a5] {\!\!\!$\lnot a$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) [left= 12.5cm of m] {\!\!\!start\!\!\!};

@@

        \end{tikzpicture}}
%        \includegraphics[scale=0.3]{examples/ex9/ex9-unsimplified.png}
        \caption{Timeline for %un-simplified 
        $\omega$-regular expression $((\lnot a) | (aa^*(\lnot a)))((\lnot a) | (aa^*(\lnot a)))^{\omega}$}
        \label{fig:unsimplified}
    \end{figure}
    \begin{figure}[h!]
        \centering
        \resizebox{.8\textwidth}{!}{
        \begin{tikzpicture}
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
  %      \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p3) [left= 1cm of p2] {\!\!\!$\lnot p$\!\!\!};
        \node (m) at ($(a1)!0.5!(a2)$) {};
        \node[shape=rectangle, fill=gray!30!white, style={minimum width=10.9cm, minimum height=3cm}] (g) [right= -1.1cm of m] {};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a1) {\!\!\!$\lnot a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a2) [below=0.4cm of a1] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a3) [right= 0.75cm of a2] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a4) [right= 0.75cm of a3] {\!\!\!$\dots$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a5) [right= 0.75cm of a4] {\!\!\!$a$\!\!\!};
        \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (a6) [right= 0.75cm of a5] {\!\!\!$\lnot a$\!\!\!};
        \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) [left= 2cm of m] {\!\!\!start\!\!\!};
        \draw[thick, line width=1.25] (s) edge[-latex] (a1) (s) edge[-latex] (a2) (a2) edge[-latex] (a3) (a3) edge[-latex] (a4) (a4) edge[-latex] (a5) (a5) edge[-latex] (a6);
        \end{tikzpicture}}
  %      \includegraphics[scale=0.3]{examples/ex9/ex9.png}
        \caption{Timeline for simplified $\omega$-regular expression $((\lnot a) | (aa^*(\lnot a)))^{\omega}$}
        \label{fig:simplified}
    \end{figure}
\end{example}

@@

\begin{definition}[Timeline Length]
    We define the timeline length of a regular expression $\tllen(A)$ recursively as follows
    \begin{align*}
        \tllen(\emptyset)   &= -\infty\\
        \tllen(\epsilon)        &= 0\\
        \tllen(a)           &= 1\\
        \tllen(A_1 A_2)     &= \tllen(A_1) + \tllen(A_2)\\
        \tllen(A_1 + A_2)   &= \max(\tllen(A_1), \tllen(A_2))\\
        \tllen(A^*)         &= 2\tllen(A) + 1
    \end{align*}
    We also extend this definition to $\omega$-regular expressions as follows
    \begin{align*}
        \tllen(A^\omega)    &= \tllen(A)\\
        \tllen(AB)          &= \tllen(A) + \tllen(B)\\
        \tllen(B_1 + B_2)   &= \max(\tllen(B_1), \tllen(B_2))
    \end{align*}
    This quantity $\tllen$ captures the length of the longest path in our timeline graph visualization.
\end{definition}
